The unequal distribution of access to education and transport: insights from the São Paulo Metropolitan Region

### 2017-02-26 paper.tex
Finalise description of Google Matrix API

Our proposed competitive accessibility measure combines information
about the place of residence of students, the spatial distribution of
public schools, and public transport accessibility into a single
measure. We start from a cumulative opportunity measure that counts the
number of schools seats that can be reached within a 30 minute journey
by public transport, and then build a competitive measure that takes
into account the place of residency of teens. Travel-to-school times by
public transport are based on actual commuting times obtained through
the Google Matrix Distance API. We calculate these measures for 633
areas within the SPMR. We use the competitive measure to simulate the
impact of a policy change in the location of public secondary schools,
in order to, first, understand the extent of spatial inequalities in
school accessibility by public transport, and second, estimate the
effect of a policy aiming at concentrating public secondary education
provision on school accessibility. We find that closing down schools in
areas with lower than average provision is highly regressive: it
negatively impacts students with the lowest accessibility levels,
accentuating existing inequalities.

>>>>

Our proposed competitive accessibility measure combines information
about the place of residence of students, the spatial distribution of
public schools, and public transport accessibility into a single
measure. We start from a cumulative opportunity measure that counts the
number of schools seats that can be reached within a 30 minute journey
by public transport, and then build a competitive measure that takes
into account the place of residency of teens. Travel-to-school times by
public transport are based on actual commuting times obtained through
use of routing algorithms provided on-line through an on-line service.

For this work we used the Google Matrix Distance API
(API stands for `application programming interface' and is used for
compter-computer communication over the web), a multi-modal
real-time travel data provider (see https://developers.google.com/maps/documentation/distance-matrix/). 
The API was used in preference to other options,
such as Routino, described in a paper by Singleton (2014),
and the OpenStreetMap Routing Machine (OSRM)
because its
ease of use (requiring no new software to be installed locally)
and performance: the Google Distance Matrix
API returns
route distance, price and times for journeys between origins and
destinations, provided either as Longitude/Latidue pairs or
as text strings to be 'geocoded' (converted to geographical location).

We calculate these measures for 633
areas within the SPMR. We use the competitive measure to simulate the
impact of a policy change in the location of public secondary schools,
in order to, first, understand the extent of spatial inequalities in
school accessibility by public transport, and second, estimate the
effect of a policy aiming at concentrating public secondary education
provision on school accessibility. We find that closing down schools in
areas with lower than average provision is highly regressive: it
negatively impacts students with the lowest accessibility levels,
accentuating existing inequalities.
